From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-07-22 dictionary.tex

\section{Quasicharacter sheaves for $p$-adic tori}\label{sec:applications}

@@

 \begin{remark}
In Section~\ref{ssec:CS_tori} we see that the \'etale site on $\Gr_R(X)$ is rich enough to geometrize all
quasicharacters of $X(K)$ as $\ell$-adic local systems on $\Gr_R(X)$, where $X$ is a N\'eron model for an
algebraic torus or an abelian variety over a local field $K$. 
It is natural to ask if the \'etale site on the generic fibre $X_K$ would have sufficed.
This seems unlikely, since the geometric \'etale fundamental group of $\Gm{K}$ is ${\hat \ZZ}$;
however, limited results in this direction were established in \cite{Cunningham-Kamgarpour} when $K = \mathbb{Q}_p$.
\end{remark}

\subsection{Quasicharacter sheaves for $p$-adic tori} \label{ssec:CS_tori}

@@

\begin{example}
We give 
\end{example}

>>>>

\section{Quasicharacter sheaves for \texorpdfstring{$p$}{p}-adic tori}\label{sec:applications}

@@

 \begin{remark}
In Section~\ref{ssec:CS_tori} we see that the \'etale site on $\Gr_R(X)$ is rich enough to geometrize all
quasicharacters of $X(K)$ as $\ell$-adic local systems on $\Gr_R(X)$, where $X$ is a N\'eron model for an
algebraic torus or an abelian variety over a local field $K$. 
It is natural to ask if the \'etale site on the generic fibre $X_K$ would have sufficed.
This seems unlikely, since the geometric \'etale fundamental group of $\Gm{K}$ is ${\hat \ZZ}$;
however, limited results in this direction were established in \cite{cunningham-kamgarpour:13a} when $K = \mathbb{Q}_p$.
\end{remark}

\subsection{Quasicharacter sheaves for \texorpdfstring{$p$}{p}-adic tori} \label{ssec:CS_tori}

@@

\begin{example}
When $T_K = \Gm{K}$ or $T_K = U_1(K'/K)$, the geometric component group $\pi_0(T)_{\bar{k}}$ is cyclic,
so $\TrFrob{\Gr_R(T)}$ is an isomorphism.  Conversely, when $T_K = \Gm{K}^2$ then
\[
\Hh^0(\Weil{\Fq}, \Hh^2(\pi_0(T), \EEx)) = \EEx,
\]
and there are uncountably many invisible quasicharacter sheaves for $T$.

We may also give examples of tori whose N\'eron models have component groups appearing in Example \ref{eg:H2}.
Let $L = K'(\sqrt{\pi})$ be a quadratic ramified extension of $K'$.  When $K = K'$ and $T_K = U_1(L/K) \times U_1(L/K)$,
the component group $\pi_0(T)_{\bar{k}}$ is $\ZZ/2\ZZ \times \ZZ/2\ZZ$ with trivial Frobenius action.  When $K'/K$ is an unramified
quadratic extension and $T_K = \Res_{K'/K} U_1(L/K')$, then $\pi_0(T)_{\bar{k}}$ is $\ZZ/2\ZZ \times \ZZ/2\ZZ$ with Frobenius
exchanging the direct factors.  Finally, let $K'/K$ be a cubic unramified extension and $S_K = \Res_{K'/K} U_1(L/K')$.  If
$T_K$ is the subtorus with character lattice $X^*(S_K) / \langle (1,1,1) \rangle$, then $\pi_0(T)_{\bar{k}}$ is $\ZZ/2\ZZ \times \ZZ/2\ZZ$
with Frobenius of order $3$.  Each of these tori will have one invisible quasicharacter sheaf.
\end{example}

We may also extract information about the automorphism groups of quasicharacter sheaves from Theorem \ref{thm:CSXK}.
